---
title: "In-Class Lab 11"
author: "ECON 425 (Justin Heflin, West Virginia University)"
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{multicol}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{xcolor}
date: "April 26, 2023"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = 'hide', fig.keep = 'none')
```

The purpose of this lab is to practice using R for using dummy variables in R. The lab may be completed as a group. To receive credit, upload your .R script to the appropriate place on eCampus (``In-Class Labs'' folder).

## For starters
Open a new R script (named `ICL11_XYZ.R`, where `XYZ` are your initials)

## Clean out/``Sweep'' R Studio
Click the broom in the Environment panel (top-right), it is directly below the Tutorial button. Also, in the bottom-right panel, click the Plots button and then click the broom in that panel. This should help with loading things into R. 

## R Packages
```
install.packages("magrittr")
install.packages("broom")
```
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(ggplot2)
library(wooldridge)
library(modelsummary)
```

## Load the data
We'll use a new data set on extramarital affairs, called `affairs` from the Wooldridge R Package.
```{r}
affair_data <- as.data.frame(affairs)
```

\textbf{Variable Names and brief description:}
\begin{itemize}
\item male: =1 if male
\item age: in years
\item yrsmarr: years married
\item kids =1 if have kids
\item ratemarr: 5 = vry hap marr, 4 = hap than avg, 3 = avg, 2 = smewht unhap, 1 = vry unhap
\item affair: =1 if had at least one affair
\item naffairs: number of affairs within last year
\end{itemize}

Check out what's in the data by typing
```{r}
summary(affair_data)
```
You'll notice that there are a number of variables that only take on 0/1 values: `male`, `kids`, `affair`, `hapavg`, `vryrel`, etc. There are also variables that take on a few different values: `relig`, `occup`, and `ratemarr`.

You can create a histogram of a factor variable in `ggplot()` as follows:
```{r}
ggplot(affair_data,aes(x=ratemarr)) + geom_bar()
```

## Multiple regression with factor variables
Let's run a regression with `naffairs` as the dependent variable and `male`, `yrsmarr`, `kids`, and `ratemarr` as the covariates.
```{r}
regression <- lm(naffairs ~ male + yrsmarr + kids + ratemarr, data=affair_data)
summary(regression)
```

## Linear Probability Model
Let's run the same regression as before, but this time use `affair` as the dependent variable. What happens when you run the following code?
```{r message=FALSE, warning=FALSE}
regression_2 <- lm(affair ~ male + yrsmarr + kids + ratemarr, data=affair_data)
summary(regression_2)
```

Finally, let’s run a more flexible model where we allow the effect of fathers and mothers to be different. The
way to do this in lm() is as follows:
```{r}
regression_3 <- lm(affair ~ male*kids + yrsmarr + ratemarr, data=affair_data)
summary(regression_3)
```