---
title: "In-Class Lab 7"
author: "ECON 425 (Justin Heflin, West Virginia University)"
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{multicol}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{xcolor}
date: "March 3, 2023"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The purpose of this lab is to practice using R to conduct hypothesis tests. The lab may be completed as a group. To receive credit, upload your .R script to the appropriate place on eCampus (``In-Class Labs'' folder).

## For starters
Open a new R script (named `ICL7_XYZ.R`, where `XYZ` are your initials)

## Clean out/``Sweep'' R Studio
Click the broom in the Environment panel (top-right), it is directly below the Tutorial button. Also, in the bottom-right panel, click the Plots button and then click the broom in that panel. This should help with loading things into R. 

## Hypothesis Testing

\begin{equation*}
colGPA = \beta_{0} + \beta_{1}hsGPA + \beta_{2}ACT + \beta_{3}skipped + \epsilon
\end{equation*}

\begin{itemize}
\item colGPA: college GPA
\item hsGPA: high school GPA
\item ACT: ACT score
\item skipped: average lectures missed per week 
\end{itemize}

Let's develop hypotheses for each slope coefficient: 


$H_{0}: \beta_{hsGPA} = 0 ; \hspace{4mm} H_{0}: \beta_{ACT} = 0 ; \hspace{4mm} H_{0}: \beta_{skipped} = 0$ 

$H_{A}: \beta_{hsGPA} \neq 0 ; \hspace{4mm} H_{A}: \beta_{ACT} \neq 0 ; \hspace{4mm} H_{A}: \beta_{skipped} \neq 0$

Each hypothesis test for slope coefficients above are two-tailed.

\begin{equation*}
\widehat{colGPA} = 1.39 + \underset{\textcolor{red}{(.094)}}{.412hsGPA} + \underset{\textcolor{red}{(.011)}}{.015ACT} - \underset{\textcolor{red}{(.026)}}{.083skipped}
\end{equation*} \vspace{2cm}


### Calculate t-statistic for each slope coefficient
\begin{multicols}{3}
\begin{equation*}
t_{hsGPA} = \frac{.412}{.094} \rightarrow 4.383
\end{equation*} \break
\begin{equation*}
t_{ACT} = \frac{.015}{.011} \rightarrow 1.364
\end{equation*} \break
\begin{equation*}
t_{skipped} = \frac{-.083}{.026} \rightarrow -3.192
\end{equation*}
\end{multicols}

Degrees of Freedom (DF) = 137 (N - K - 1 $\Longrightarrow$ 141 - 3 - 1) 

Select level of significance: $\alpha$ = 5\% (0.05)
\begin{itemize}
\item Critical t-value based on the DF, chosen level of significance, and two-tail test is about 1.96 from t-table
\end{itemize}

Compare $t_{skipped}$ = -3.192 to critical t-value $t_c$ = 1.96 using the decision rule:

(Reject $H_{0}$ if $|t_k| > t_{c}$ AND if $t_k$ also has the sign implied by $H_{A}$)

$\Longrightarrow |-3.192| > 1.96 \hspace{2mm}; \hspace{1mm} \beta_{skipped} \neq 0$ which matches our $H_{A}$  

This means we can reject the null hypothesis can conclude skipping lectures does indeed tend to have a negative relationship with college GPA (holding the other variables in the equation constant)

## Hypothesis Testing in R  
```{r}
library(wooldridge)
GPA_data <- as.data.frame(gpa1)
```


```{r}
regression1 <- lm(colGPA ~ hsGPA + ACT + skipped, data = GPA_data)
summary(regression1)
```

Now let's take a look at the confidence interval for each regression coefficient
```{r}
confint(regression1, level = 0.95)
```

```{}
install.packages("GGally")
```

```{r}
library(ggplot2)
library(GGally)
ggcoef(regression1, exclude_intercept = TRUE, vline_linetype = "solid",
       vline_color = "blue", errorbar_color = "darkorange",
       errorbar_height = 0.15)
```
