---
title: "In-Class Lab 10"
author: "ECON 425 (Justin Heflin, West Virginia University)"
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{multicol}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{xcolor}
date: "April 14, 2023"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = 'hide', fig.keep = 'none')
```

The purpose of this lab is to practice using R for time series models. The lab may be completed as a group. To receive credit, upload your .R script to the appropriate place on eCampus (``In-Class Labs'' folder).

## For starters
Open a new R script (named `ICL10_XYZ.R`, where `XYZ` are your initials)

## Clean out/``Sweep'' R Studio
Click the broom in the Environment panel (top-right), it is directly below the Tutorial button. Also, in the bottom-right panel, click the Plots button and then click the broom in that panel. This should help with loading things into R. 

## R Packages
```
install.packages("dplyr")
```
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(ggplot2)
library(wooldridge)
library(dplyr)
library(modelsummary)
library(sandwich)
library(lmtest)
```

## Load the data
We are going to use the `intdef` data-set from the `Wooldridge` R Package
```{r}
macro_data <- as.data.frame(intdef)
summary(macro_data)
```

This data-set is time-series with annual observations from 1948 to 2003 (56 observations) on 13 variables.

\textbf{Variable Names and brief description:}
\begin{itemize}
\item i3: 3 month Treasury bill rate
\item inf: CPI inflation rate
\item rec: federal receipts, percent GDP
\item out: federal outlays, percent GDP
\item def: out - rec
\end{itemize}

## Plot time series data
Let's have a look at the inflation rate for the US over the period 1948-2003:
```{r}
ggplot(macro_data, aes(year, inf)) + geom_line()
```

## Determinants of interest rate using a Distributed Lag Model 

Now let's estimate the following regression model:
\begin{equation*}
i3_{t} = \beta_0 + \beta_{1}inf_{t} + \beta_{2}inf_{t-1} + \beta_{3}inf_{t-2} + \beta_{4}def_{t} + u_{t}
\end{equation*}

```{r}
regression <- lm(i3 ~ inf + lag(inf,1) + lag(inf,2) + def, data = macro_data)
summary(regression)
```

## Using Newey-West Standard Errors on Distributed Lag Model
```{r}
NW_VCOV <- NeweyWest(regression)
coeftest(regression, vcov. = NW_VCOV)
```

## Determinants of interest rate using a Lagged Dependent Variable Model

\begin{equation*}
i3_{t} = \beta_0 + \beta_{1}inf_{t} + \beta_{2}i3_{t-1} + \beta_{3}i3_{t-2} + \beta_{4}def_{t} + u_{t}
\end{equation*}

```{r}
regression_2 <- lm(i3 ~ inf + lag(i3, 1) + lag(i3,2) + def, data = macro_data)
summary(regression_2)
```

## Using Newey-West Standard Errors on Lagged Dependent Variable Model
```{r}
NW_VCOV_2 <- NeweyWest(regression_2)
coeftest(regression_2, vcov. = NW_VCOV_2)
```

## Determinants of interest rate using both classes of Dynamic Models

\begin{equation*}
i3_{t} = \beta_0 + \beta_{1}inf_{t} + \beta_{2}inf_{t-1} + \beta_{3}i3_{t-1} + \beta_{4}def_{t} + u_{t}
\end{equation*}

```{r}
regression_3 <- lm(i3 ~ inf + lag(inf,1) + lag(i3,1) + def, data = macro_data)
summary(regression_3)
```

## Using Newey-Standard Errors on Dynamic Model
```{r}
NW_VCOV_3 <- NeweyWest(regression_3)
coeftest(regression_3, vcov. = NW_VCOV_3)
```




